\section{Research Plan and Timeline}
\label{sec:timeline}

I have already made progress towards my thesis:

\begin{itemize}
\item I have created the formalism for describing the operational
  semantics of linklets.
\item I have implemented the linklets in RPython and incorporated them into
  Pycket.
\item I have implemented all the run-time support and primitives to
  incorporate the expander into Pycket and be able to implement
  \texttt{\#racket}.
\item I have created a formalism for a stackful and a CEK-based
  interpreter working together.
\item I have prototyped the two interpreter approach where a stackful
  interpreter is working along with the existing CEK that is hosting
  Racket using linklets.
\end{itemize}

To complete my thesis, I plan to follow this timeline:

\begin{itemize}
  \item \textbf{[December 2019]} Improve the stackful+CEK Pycket.
    \begin{itemize}
    \item Improve the stackful interpreter enough to run more complex
      programs with linklets.
     \item Investigate and improve the strategies for switching with
       the CEK.
     \item Devise and implement experiments to show that it is a
       viable approach by demonstrating that it helps with the GC
       issue.
    \end{itemize}
  \item \textbf{[January 2018]} Improve the performance of regexp
    implementation.
    \begin{itemize}
      \item Devise and implement strategies to get the performance of
        the regexp implementation closer to the existing RPython
        regexp.
      \item Plan and execute experiments to demonstrate that it is a
        viable approach to meta-trace programs with complex control
        flow.
    \end{itemize}
  \item \textbf{[February-May 2020]} Write dissertation
  \item \textbf{[June 2020]} Defend
\end{itemize}

\subsection{Planned Publications}

I plan to publish the following papers:

\begin{itemize}
\item \textbf{2020} \emph{Best of both worlds: Two interpreters working together.} Which conference? When?
\item \textbf{2020} \emph{When JIT hits the fan: Meta-tracing complex control flows.} Which conference? When?
\end{itemize}
